---
title: "p8105_hw6_zz2780"
author: "Lydia Zhang"
date: "2022-12-01"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 2

Load the data from Github Repo
```{r}
data2<-read.csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")
```

The code below creates a city_state variable and a binary variable solved indicating whether the homicide is solved, omits cities with missing race or mistake, limits the victim's race to only white or black, and changes victim_age to numeric variable. 
```{r}
data2<-data2%>%
  mutate(
    city_state=paste(city, state, sep=", "),
    solved=if_else(str_starts(disposition, "Closed"),1,0),
    victim_age=as.numeric(victim_age))%>%
  filter(
   !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
   victim_race %in% c("White", "Black")
  )
```

After subset the dataset to only for city of Baltimore, I ran a glm on the dataset with solved/unsolved homocide as outcome, and victim's age, sex, and race as predictors.
```{r}
data2_baltimore=data2%>%
  filter(
    city_state=="Baltimore, MD"
  )

fit=glm(solved ~ victim_age + victim_sex + victim_race, data=data2_baltimore)
```

The code below tidy the glm output, and obtains the odds ratio and its 95% confidence interval for solving homicides comparing male victims to female victims keeping all other variables fixed.
```{r}
fit%>%
  broom::tidy()%>%
  select(term, estimate, std.error)%>%
  filter(term=="victim_sexMale")%>%
  mutate(
    odds_ratio=exp(estimate),
    lower_ci=exp(estimate-1.96*std.error),
    upper_ci=exp(estimate+1.96*std.error)
  )%>%
  select(-estimate, -std.error)
```
The code below runs glm for every city in the data set and saves the glm output in results, which is then unnested.
```{r}
nest_data2<-data2%>%
  nest(data=-city_state)%>%
  mutate(
    models=purrr::map(data,~glm(solved ~ victim_age + victim_sex + victim_race, data=.x)),
    results=purrr::map(models, broom::tidy)
  )%>%
  select(-data, -models)%>%
  unnest(results)
```

The code below creates a dataframe odds_ratio_sex that stores the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims for every city, keeping all other variables fixed
```{r}
odds_ratio_sex<-nest_data2%>%
  select(city_state, term, estimate, std.error)%>%
  filter(term=="victim_sexMale")%>%
  mutate(
    odds_ratio=exp(estimate),
    lower_ci=exp(estimate-1.96*std.error),
    upper_ci=exp(estimate+1.96*std.error)
  )%>%
  select(-term, -estimate, -std.error)

odds_ratio_sex
```
The barplot plot_or shows the estimated ORs and CIs for solving homicides comparing male and female victims while keeping other variables constant, in descending order of OR. The ORs are represented by the bars and CIs are represented by the error bars. 

We can observe that Stockton, CA has OR closest to 1, indicating that solving homicides in Stockton, CA does not associate with victim's gender. Only Fresno, CA, Minneapolis, MN, and Stockton, CA have OR greater than 1, meaning that all other cities would be more likely to solve the homicide if the victim is female, with New York City has the lowest OR. 
```{r}
plot_or<- ggplot(odds_ratio_sex, aes(x =reorder(city_state, desc(odds_ratio)), y = odds_ratio)) +
  geom_bar(stat="identity")+
  geom_errorbar(width=.1, aes(ymin=lower_ci,ymax=upper_ci),position=position_dodge(width=0.5))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(x="City, State", y = "Odds Ratio",title="Estimated Odds Ratio and CI for Solving Homicides Comparing Male victims to Female victims in Each City")

plot_or
```


Question 3

The code below loads the data, clean the variable names, and covert numeric variables to factor where appropriate.
```{r}
data3<-read_csv("birthweight.csv")%>%
  janitor::clean_names()%>%
  mutate(
    babysex<-as.factor(babysex),
    frace<-as.factor(frace),
    malform<-as.factor(malform),
    mrace<-as.factor(mrace)
  )
```
I used complete.cases to check if data 3 has any row that have missing values. It returns none, meaning the data has no missing value.
```{r}
data3[!complete.cases(data3),]
```

I hypothesize that the mom's pre-pregnancy BMI is positively associated with baby's birthweight, and I think the presence of malformation can be an effect modifier (EMM) in the regression model (model1). I fitted a linear regression model with mother's BMI, presencence of malformations, and their interaction term as predictors and baby's birthweight as outcome.
```{r}
model1<-lm(bwt~ppbmi*malform, data=data3)
```

I added residuals and fitted values of the model and created a violin plot of model residuals against fitted value. 
```{r}
data3 %>% 
  modelr::add_residuals(model1) %>% 
  modelr::add_predictions(model1)%>%
  ggplot(aes(x = resid, y = pred)) + geom_violin()
```




